\documentclass[11pt]{article}
\usepackage{ndstyle}
\newcommand{\checkbox}{\fbox{\rule{0pt}{1.2ex}\rule{1.2ex}{0pt}}}
\setcounter{secnumdepth}{0}
\setcounter{tocdepth}{0}

\title{New Doggerland — Governing Body Vetting Checklist}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Status}

\noindent
\textbf{Status:} Non-binding checklist \\
\textbf{Authority:} Derived from \emph{New Doggerland — Board Vetting Process} \\
\textbf{Purpose:} Provide a concise, practical record of constitutional eligibility screening

This checklist documents vetting steps only. It does not grant authority, waive requirements, or override any governing document.

\vspace{1em}

\section*{Candidate Information}

\begin{itemize}
\item Candidate Name:\hspace{4em}\rule{10cm}{0.4pt}
\item Date of Vetting:\hspace{4.5em}\rule{10cm}{0.4pt}
\item Vetting Conducted By:\hspace{1.5em}\rule{10cm}{0.4pt}
\end{itemize}

\vspace{0.5em}

\section*{Stage 1 — Independence and Disclosure}

\begin{itemize}
\item[\checkbox] Candidate affirms ability to act independently.
\item[\checkbox] Candidate is not the Founder.
\item[\checkbox] Candidate is not the Founder’s spouse.
\item[\checkbox] Candidate is not a member of the Founder’s household.
\item[\checkbox] Candidate has disclosed all Material Financial Interests.
\item[\checkbox] No undisclosed prohibited relationships identified.
\item[\checkbox] Candidate acknowledges ongoing disclosure obligation.
\end{itemize}

\noindent
Stage 1 Result: \hspace{1em} \checkbox\ Pass \hspace{2em} \checkbox\ Fail

\vspace{1em}

\section*{Stage 2 — Welfare Commitment}

Candidate demonstrates a sustained, good-faith commitment to one or more of the following:

\begin{itemize}
\item[\checkbox] Canine or animal welfare
\item[\checkbox] Human welfare
\item[\checkbox] Responsible animal stewardship
\end{itemize}

Evidence type(s) reviewed:

\begin{itemize}
\item[\checkbox] Long-term dog ownership or caregiving
\item[\checkbox] Foster, rescue, or shelter involvement
\item[\checkbox] Veterinary, behavioral, or welfare-adjacent professional work
\item[\checkbox] Governance, financial, or operational support of welfare organizations
\item[\checkbox] Sustained volunteer or advocacy work
\item[\checkbox] Other documented conduct (described below)
\end{itemize}

Evidence notes / references:

\vspace{1.5em}

\noindent
Stage 2 Result: \hspace{1em} \checkbox\ Pass \hspace{2em} \checkbox\ Fail

\vspace{1em}

\section*{Stage 3 — Disqualification Screening}

\begin{itemize}
\item[\checkbox] No conviction for animal cruelty.
\item[\checkbox] No conviction for animal abuse.
\item[\checkbox] No conviction for animal neglect.
\item[\checkbox] No substantiated judicial or regulatory finding for animal cruelty, abuse, or neglect.
\item[\checkbox] No conviction or substantiated finding for cruelty toward humans.
\end{itemize}

Reasonable confirmation relied upon:

\begin{itemize}
\item[\checkbox] Candidate written attestation
\item[\checkbox] Public records review
\item[\checkbox] Other information reasonably available at the time
\end{itemize}

\noindent
Stage 3 Result: \hspace{1em} \checkbox\ Pass \hspace{2em} \checkbox\ Fail

\vspace{1em}

\section*{Stage 4 — Continuing-Condition Acknowledgment}

\begin{itemize}
\item[\checkbox] Candidate acknowledges ongoing compliance requirement.
\item[\checkbox] Candidate acknowledges mandatory review and removal upon discovery of disqualifying conduct.
\item[\checkbox] Candidate acknowledges omission or concealment may itself constitute cause for removal.
\end{itemize}

\vspace{1em}

\section*{Final Determination}

\begin{itemize}
\item[\checkbox] Eligible for appointment
\item[\checkbox] Not eligible for appointment
\end{itemize}

\vspace{1em}

\noindent
Reviewer Signature:\hspace{3em}\rule{6cm}{0.4pt}\hspace{2em}Date:\rule{3cm}{0.4pt}

\vspace{1em}

\noindent
This checklist must be retained in governance records. Failure to complete this checklist does not validate an ineligible appointment.

\end{document}
