\documentclass[11pt]{article}
\usepackage{ndstyle}
\setcounter{secnumdepth}{1}

\title{\textbf{New Doggerland}\\
\vspace{0.25em}
\large Donor-Advised Fund (DAF) Safety \& Accountability Overview}
\author{}
\date{}

\begin{document}
\maketitle

\noindent
\textbf{Status:} External explanatory overview (non-binding) \\
\textbf{Authority:} Subordinate to all governance and operational documents \\
\textbf{Contact for compliance inquiries:} \texttt{compliance@newdoggerland.org}

\vspace{1em}

\section*{Purpose of this document}

This document explains how donor funds are \textbf{constrained, monitored, and reviewable}
within New Doggerland’s governance and operations framework.
It is written for Donor-Advised Fund (DAF) administrators, compliance teams, and donors
seeking \textbf{high-confidence, low-surprise grant pathways}.

This document does not create authority, permissions, or obligations.
If any discrepancy exists, controlling governance and operational documents prevail.

\section*{Rationale for Early-Stage Funding}

Donor-advised funds and other fiduciary intermediaries are typically advised to prioritize organizations with established operating history, scale, and demonstrated outcomes. This guidance reflects the reality that governance and accountability systems are often developed reactively, after growth has begun.

New Doggerland was intentionally designed in the opposite order.

Governance, control, and accountability mechanisms were established prior to scale, reputation, or external validation. This allows funders to evaluate safety and stewardship directly, rather than inferring them from success.

As organizations mature, governance systems tend to harden around existing habits and legacy decisions. Early-stage formation is therefore the only period in which safety can be designed without exception or retrofitting.

For this reason, early funding of New Doggerland is not a speculative commitment to outcomes. It is a bounded opportunity to observe and validate governance functioning under real but constrained operating conditions.

\subsection*{Role of Early Funders}

Early funders support operations within predefined thresholds and observe governance processes such as escalation, logging, and board engagement as they occur.

Early funders do not underwrite rapid expansion, discretionary growth, or outcome-driven urgency. Their role is to validate that governance systems operate as designed before scale introduces additional complexity.


\section*{How donor funds are protected by design}

New Doggerland is structured so that \textbf{no individual can unilaterally control donor funds},
including founders, staff, or emergency actors.

Key protections include:
\begin{itemize}
    \item Explicit \textbf{Material Action thresholds} requiring Governing Body approval for spending, contracts, compensation, and access.
    \item \textbf{Aggregation rules} preventing incremental bypass of approval thresholds.
    \item \textbf{Segregation of duties} prohibiting the same person from initiating and approving financial actions.
    \item \textbf{Emergency authority} that is explicitly barred from affecting donor funds, contracts, or compensation.
\end{itemize}

Any financial action must be authorized at the appropriate level, logged at the time of action,
and subject to post-action review.

\section*{Where donor funds can and cannot go}

Subject to donor restriction and approval level, funds may be used for:
\begin{itemize}
    \item Direct animal welfare operations and care
    \item Safety and infrastructure improvements
    \item Emergency veterinary response (non-compensatory)
    \item Governance, audit, and compliance capacity
\end{itemize}

Donor funds may not be used for:
\begin{itemize}
    \item Founder or staff compensation without explicit approval
    \item Expansion beyond operational capacity
    \item Contracts or obligations created under emergency authority
    \item Informal, undocumented, or retrospective actions
    \item Any purpose outside New Doggerland’s mission lock
\end{itemize}

All uses are evaluated against welfare, dignity, and long-horizon stewardship constraints.

\section*{Emergency situations do not bypass accountability}

New Doggerland is designed to operate safely under stress.

\begin{itemize}
    \item Emergency authority is narrowly scoped, time-limited, and logged.
    \item Emergency actions do not establish precedent.
    \item Repeated or extended emergency use automatically triggers review.
    \item Emergency authority cannot access or move donor funds.
\end{itemize}

Emergencies are treated as reviewable events, not justifications for discretion.

\section*{Logging, auditability, and transparency}

All material actions are subject to a \textbf{tamper-evident, append-only logging system}:

\begin{itemize}
    \item Logs are created at the time of action or immediately upon stabilization.
    \item Corrections occur by addition, never deletion.
    \item Access and edits are traceable.
    \item Gaps or delays automatically trigger review.
\end{itemize}

This structure supports internal audit, external audit, donor inquiry, and regulator inquiry.

\section*{Governance resilience and founder constraints}

\begin{itemize}
    \item The Founder holds no unilateral authority over funds or policy.
    \item The Founder cannot self-authorize compensation or benefit.
    \item Governance continues independently of any one individual.
    \item Conflicts of interest are screened and documented.
\end{itemize}

The system is designed to outlive personalities and remain legible to external reviewers.

\section*{Safe exit and corrective pathways}

DAF safety includes the ability to pause, redirect, or unwind.

New Doggerland’s framework supports:
\begin{itemize}
    \item Suspension of activities that exceed capacity
    \item Corrective action following logged failure
    \item Donor-restricted fund accounting
    \item Clean project wind-down if required
\end{itemize}

There are no informal override paths.

\section*{Why this matters for DAFs}

Most organizations rely on trust.
New Doggerland is designed for \textbf{scrutiny}.

\begin{itemize}
    \item Constraints are explicit.
    \item Authority is bounded.
    \item Failures are legible.
    \item Emergencies do not expand power.
    \item Oversight is structurally invited.
\end{itemize}

This makes New Doggerland a suitable recipient for restricted, high-accountability donor funds.

\end{document}
